---
title: "parafac improvements"
output: html_document
  toc: true
  toc_depth: 2
---

# preface

Goals:

- More robust PARAFAC (con: more computationally intensive)

- Physically reasonable results

# libraries

```{r}
library(staRdom)
library(tidyverse)
```

# model 0
## data prep 0.0

```{r}

eems <- eem_read_csv(path = "../Data/corrected csvs 190221", 
                     is_blank_corrected = TRUE,
                     is_scatter_corrected = TRUE,
                     is_ife_corrected = TRUE,
                     is_raman_normalized = TRUE,
                     manufacturer = "Horiba")

eems_smp <- eems[sapply(eems, function(x) str_detect(x$sample, "SMP"))]

class(eems_smp) <- "eemlist"

eems_em <- eems[sapply(eems, function(x) str_detect(x$sample, "EM"))]

class(eems_em) <- "eemlist"

```

## model run 0.0

```{r}

pf_0_all <- eem_parafac(eems, comps = 4:10, maxit = 5000, normalise = TRUE,
                        nstart = 15, ctol = 2e-5)

pf_0_non <- eem_parafac(eems, comps = 4:10, maxit = 5000, normalise = FALSE,
                        nstart = 15, ctol = 2e-5)

pf_0_em <- eem_parafac(eems_em, comps = 4:10, maxit = 5000, normalise = TRUE,
                       nstart = 15, ctol = 2e-5)

```

## model comparison 0.0

```{r}

eempf_compare(pf_0_all)

eempf_compare(pf_0_non)

eempf_compare(pf_0_em)

```

## model truthing 0.0

```{r}

eempf_corplot(pf_0_all[[5]]) ##8 comp.

eempf_corplot(pf_0_non[[1]]) ##4 comp.

eempf_corplot(pf_0_em[[4]])  ##7 comp.

```

## outlier selection 0.0

```{r}

all8lev <- eempf_leverage(pf_0_all[[5]])

non4lev <- eempf_leverage(pf_0_non[[1]])

em7lev <- eempf_leverage(pf_0_em[[4]])

eempf_leverage_plot(all8lev, qlabel = .1)

eempf_leverage_plot(non4lev, qlabel = .1)

eempf_leverage_plot(em7lev, qlabel = .1)

```

## outlier removal 0.0

```{r}

eems_all8 <- eem_exclude(eems, exclude = list("em" = c(541.27, 301.36, 319.063, 
                                                       317.451),
                                              "ex" = c(246, 249),
                                              "sample" = c("1604_FOXC_EMP_2", 
                                                           "1607_SFPR_EMR_1")
                                              )
                         )

eems_non4 <- eem_exclude(eems, exclude = list("sample" = c("1605_SFMD_EML_1",
                                                           "1608_ELDC_EMP_1",
                                                           "1605_SFSI_EML_1",
                                                           "1605_FOXC_EML_1",
                                                           "1605_REDC_EML_1")
                                              )
                         )

eems_em7 <- eem_exclude(eems_em, exclude = list("em" = 541.27,
                                                "ex" = c(246, 249),
                                                "sample" = c("1607_SFMG_EMY_1",
                                                             "1607_SFPR_EMR_1",
                                                             "1605_SFMD_EML_1")
                                                )
                        )

```

## model run 0.1

```{r}

pf_1_all <- eem_parafac(eems_all8, comps = 4:10, normalise = TRUE)

pf_1_non <- eem_parafac(eems_non4, comps = 4:10, normalise = FALSE)

pf_1_em <- eem_parafac(eems_em7, comps = 4:10, normalise = TRUE)

```

## model comparison 0.1

```{r}

eempf_compare(pf_1_all)

eempf_compare(pf_1_non)

eempf_compare(pf_1_em)

```

## model truthing 0.1

```{r}

eempf_corplot(pf_1_all[[5]]) ##8 comp.

eempf_corplot(pf_1_non[[1]]) ##4 comp.

eempf_corplot(pf_1_em[[3]])  ##6 comp.

```

## outlier selection 0.1

```{r}

all8lev <- eempf_leverage(pf_1_all[[5]])

non4lev <- eempf_leverage(pf_1_non[[1]])

em6lev <- eempf_leverage(pf_1_em[[3]])

eempf_leverage_plot(all8lev, qlabel = .1)

eempf_leverage_plot(non4lev, qlabel = .1)

eempf_leverage_plot(em6lev, qlabel = .1)

```

## outlier removal 0.1

```{r}

eems_all8 <- eem_exclude(eems_all8, 
                         exclude = list("ex" = c(252, 255),
                                        "sample" = c("1607_SFMG_EMY_1")))
                      
eems_non4 <- eem_exclude(eems_non4, 
                         exclude = list("sample" = c("1605_SFMG_EML_1",
                                                     "1608_FOXC_EMP_1",
                                                     "1608_SFMD_EML_1",
                                                     "1608_SFMG_EML_1")))
               
eems_em6 <- eem_exclude(eems_em7, exclude = list("ex" = c(252, 255)))
                        
```

## model run 0.2

```{r}

pf_2_all <- eem_parafac(eems_all8, comps = 4:10, normalise = TRUE)

pf_2_non <- eem_parafac(eems_non4, comps = 4:10, normalise = FALSE)

pf_2_em <- eem_parafac(eems_em6, comps = 4:10, normalise = TRUE)

```

## model comparison 0.2

```{r fig.width= 6, fig.height = 10}

eempf_compare(pf_2_all[3])

```

## residual analysis 0.2

```{r}

resids <- eempf_residuals(pf_2_all[[3]], eems)

top_10 <- resids %>%
  group_by(Sample) %>%
  summarise(avg_resid = mean(abs(value), na.rm = TRUE)) %>%
  arrange(desc(avg_resid)) %>%
  .[1:10,] %>%
  select(Sample) %>%
  .$Sample

outliers <- eem_extract(eems, top_10, keep = TRUE)

eempf_residuals_plot(pf_2_all[[3]], outliers, residuals_only = TRUE)

```

## output model 0.2 for openfluor analysis

```{r}

best_model <- eempf_rescaleBC(pf_2_all[[3]])

eempf_openfluor(best_model, "best_model_3_11.txt")

```




## save model 0.2

```{r}

save(pf_2_all, file = "best_parafac_3_6.RData")

```


# model 1

## data prep 1.0

Data being loaded have Component 6 from Model 0.2 trimmed.

```{r}

eems <- eem_read_csv(path = "../Data/corrected csvs 190311", 
                     is_blank_corrected = TRUE,
                     is_scatter_corrected = TRUE,
                     is_ife_corrected = TRUE,
                     is_raman_normalized = TRUE,
                     manufacturer = "Horiba")

```

